---
title: "GCReW Symposium Porewater Data Project"
author: "Zoe Read"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    number_sections: true

output_dir: "To Be Reviewed/PDF"
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(ggpubr)
library(stringr)
library(readxl)
library(tidyr)
library(tibble)
library(openxlsx)
library(knitr)
library(curl)
library(ggpmisc)
library(arrow)
library(lubridate)
library(rstatix)

```

##Grab Compiled Porewater Data Files
```{r Pull in 2022-2024 data files, message=FALSE, include=FALSE}

##SO4/Cl Data
so4_cl_data <- read.csv(curl("https://raw.githubusercontent.com/COMPASS-DOE/synoptic-discrete-data/refs/heads/main/Porewater/Sulfate_Chloride/Synoptic_CB/COMPASS_SynopticCB_SO4_Cl_AllData.csv"))

##H2S Data
h2s_data <- read.csv(curl("https://raw.githubusercontent.com/COMPASS-DOE/synoptic-discrete-data/refs/heads/main/Porewater/Sulfide/Synoptic_CB/COMPASS_SynopticCB_Sulfide_AllData.csv"))

##DOC Data
doc_data <- read.csv(curl("https://raw.githubusercontent.com/COMPASS-DOE/synoptic-discrete-data/refs/heads/main/Porewater/DOC/Synoptic-CB/COMPASS_SynopticCB_PW_DOC_AllData.csv"))

```


##Plot All SO4 data
```{r}
data_plotting_so4_cl <- so4_cl_data 

head(data_plotting_so4_cl)

data_plotting_so4_cl <- data_plotting_so4_cl %>% 
    mutate(
    Zone_1 = case_when(
      Zone == "UP" ~ "Upland",
      Zone == "SWAMP" ~ "Swamp",
      Zone == "TR" ~ "Transition",
      Zone == "WC" ~ "Wetland",
      Zone == "SW" ~ "Surface Water",
      TRUE ~ Zone  # keep everything else unchanged
    )
  ) %>%
  mutate(Zone_1 = factor(Zone_1, levels = c("Upland", "Swamp", "Transition", "Wetland", "Surface Water")))

data_plotting_so4_cl <- data_plotting_so4_cl %>% 
    mutate(
    Site_1 = case_when(
      Site == "SWH" ~ "Sweethall",
      Site == "GCW" ~ "GCReW",
      Site == "MSM" ~ "Moneystump",
      Site == "GWI" ~ "Goodwin",
      TRUE ~ Site  # keep everything else unchanged
    )
  ) %>%
  mutate(Site_1 = factor(Site_1, levels = c("Sweethall", "GCReW", "Moneystump", "Goodwin")))


#Plot samples to get a first look at concentrations (sanity check)
#group the data for plotting
data_plotting_so4_cl <- data_plotting_so4_cl %>%
  group_by(Site_1, Zone_1) %>%
  mutate(row_num = factor(row_number())) %>%  # create row_num as factor per group
  ungroup()


viz_so4_plot <-  ggplot(data_plotting_so4_cl, aes(x = as.character(Year), y = SO4_mM, fill = Zone_1)) +
  geom_bar(stat = "identity", position = position_dodge2(preserve = "total")) +
  facet_grid( ~ Site_1, scales="free") +
  scale_fill_manual(values = c(
    "Upland" = "#20063B",
    "Transition" = "#FFBC42",
    "Swamp" = "darkgrey",
    "Wetland" = "#419973",
    "Surface Water" = "#25ABE6"
  )) +
  theme_classic() +
  labs(x = " ", y = "Sulfate (mM)", title = paste0("Total Samples: ", nrow(so4_cl_data))) +
  theme(
    panel.spacing = unit(.05, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1))+ 
  theme(text = element_text(size = 25))+ 
  labs(fill="Zone")

viz_so4_plot

# Save as a PNG
png("COMPASS_SynopticCB_SO4_n.png", width = 3000, height = 1000, res = 150)
print(viz_so4_plot)  # For ggplot objects, use print()
dev.off()

```



##Read in water level data
```{r}

#download the L1 dataset from google drive
# setwd("S:/Biogeochemistry/COMPASS/L2 Data")

##Read in L2 Sensor Data from S Drive and only save necessary data
# Path to zip file
# L2 <- "S:/Biogeochemistry/COMPASS/L2 Data/v2-1"


# Find all files recursively
files <- list.files(path = "Raw Data/Water Level/", full.names = TRUE, recursive = TRUE)
# Read and combine
data_list <- lapply(files, read_parquet)
temps <- do.call(rbind, data_list)


# # List CSV files inside the zip
# files <- list.files(L2)
# files <- files[grepl("\\.parquet$", files)]
# 
# results <- list()
# 
# for (f in files) {
#   message(f)
# 
#   selected_research_names <- c("wl")
# 
#   # Read CSV directly from zip
#   dat <- read_parquet((L2, f))
# 
#   temp <- subset(dat, Instrument == "AQ600")
# 
#   gw <- temp %>%
#     filter(research_name %in% selected_research_names) %>%
#     select(TIMESTAMP, research_name, Value, F_OOB, F_OOS, Plot, Site)
# 
#   results[[f]] <- gw
# }

# #we need to bind these together
# dat <- do.call(rbind, results)
# rownames(dat) <- NULL
# dat1 <- dat %>% filter(!(F_OOB == 1 & F_OOS == 1))
# head(dat1)

# #convert to wide format
# temps <- pivot_wider(dat1, id_cols = c("TIMESTAMP", "Plot", "Site"), names_from = research_name, values_from = Value )
# head(temps)
# 
# write.csv(temps, file="Raw Data/L1_temps.csv")
# 
# 
# ##Read in raw temperature data
# files <- dir(path = "Raw Data/", pattern = "_wx-tempavg15.*\\.csv$", full.names = TRUE)
# results <- lapply(files, read.csv)
# dat <- do.call(rbind, results)
# rownames(dat) <- NULL
# dat <- dat %>%
#   select(TIMESTAMP, research_name, Value, F_OOB, F_OOS, Plot, Site)
# dat1 <- dat %>% filter((F_OOB != 1 & F_OOS != 1))
# head(dat1)

#convert to wide format
# temps <- pivot_wider(dat1, id_cols = c("TIMESTAMP", "Plot", "Site"), names_from = research_name, values_from = Value )
# head(temps)
```


## Match up the L1 temp data and add it to the metadata dataframe
```{r}
#Going to need to use site and timestamp in the metadata in order to find a date 

#first we need to make a timestamp column in metadat so that we can 
so4_cl_data <- so4_cl_data %>%
  mutate(
    Time = na_if(Time, ""),
    Time = replace_na(Time, "12:00"), 
    Time = paste0(Time, ":00"))

so4_cl_data$Date <- ifelse(so4_cl_data$Month < 10, paste0(so4_cl_data$Year, "0", so4_cl_data$Month), paste0(so4_cl_data$Year, so4_cl_data$Month))

so4_cl_data$Date_day <- ifelse(so4_cl_data$Day < 10, paste0(so4_cl_data$Date, "0", so4_cl_data$Day), paste0(so4_cl_data$Date, so4_cl_data$Day))

so4_cl_data$TIMESTAMP <- ymd_hms(paste(so4_cl_data$Date_day, so4_cl_data$Time))

#make sure time stamps are in the correct format 
so4_cl_data$TIMESTAMP <- as.POSIXct(so4_cl_data$TIMESTAMP, format = "%Y-%m-%d %H:%M:%OS")
temps$TIMESTAMP <- as.POSIXct(temps$TIMESTAMP, format = "%Y-%m-%d %H:%M:%OS")

#time rounding step 
so4_cl_data$DateTime_round <- round_date(so4_cl_data$TIMESTAMP, "15 mins")
temps$TIMESTAMP_round <- round_date(temps$TIMESTAMP, "15 mins")
  
#ensure there is only one data observation per hour 
temps_clean <-   temps %>% 
    group_by(TIMESTAMP_round, Site, Plot) %>%
    summarize(Water_level = mean(Value_MAC)) 

#join them 
so4_cl_data_L2temp <- so4_cl_data_L2temp %>% 
  left_join(temps_clean, by=c("Site"="Site", "DateTime_round"="TIMESTAMP_round", "Zone" = "Plot"))

write.csv(so4_cl_data_L2temp, file = "SO4_L2_Data.csv")

```


##Plot water level vs SO4/Cl
```{r}

so4_wl <- so4_cl_data_L2temp %>%
  filter(!str_detect(Site, "GWI"))%>%
  filter(!str_detect(Site, "SWH"))%>%
  filter(str_detect(Zone, "TR"))%>%
  filter(str_detect(Depth_cm, "10"))

##Different lines for each zone (UP, TR, WC)
viz_so4_wl_plot <- ggplot(so4_wl, aes(x = Value, y = SO4_ppm)) +
  geom_point(aes(color = Zone), size = 3) +
  stat_poly_eq(use_label(c("R2", "P")), size = 6) +
  stat_poly_line(formula = y ~ x, color = "black")+
  facet_wrap( ~ Site, scales="free", nrow = 1) +
  scale_color_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "SWAMP" = "darkgrey",
    "WC" = "#419973",
    "SW" = "#25ABE6"
  )) +
  theme_classic() +
  labs(x = "Soil VWC", y = "SO4 (ppm)", title = "Plot of soil VWC vs SO4 at 10 cm depth") +
  # theme(
  #   panel.spacing = unit(.05, "lines"), 
  #   panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) + 
  theme(text = element_text(size = 30)) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

 viz_so4_wl_plot
 

##Different lines for each zone (UP, TR, WC)
viz_cl_wl_plot <- ggplot(so4_wl, aes(x = Value, y = Cl_ppm)) +
  geom_point(aes(color = Zone), size = 3) +
  stat_poly_eq(use_label(c("R2", "P")), size = 6) +
  stat_poly_line(formula = y ~ x, color = "black")+
  facet_wrap( ~ Site, scales="free", nrow = 1) +
  scale_color_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "SWAMP" = "darkgrey",
    "WC" = "#419973",
    "SW" = "#25ABE6"
  )) +
  theme_classic() +
  labs(x = "Soil VWC", y = "Cl (ppm)", title = "Plot of soil VWC vs Cl at 10 cm depth") +
  # theme(
  #   panel.spacing = unit(.05, "lines"), 
  #   panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) + 
  theme(text = element_text(size = 30)) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

 viz_cl_wl_plot


# Save as a PNG
png("Soil_VWC_vs_SO4_TR.png", width = 2000, height = 1200, res = 100) 
print(viz_so4_wl_plot)  # For ggplot objVWCts, use print()
dev.off() 

# Save as a PNG
png("Soil_VWC_vs_Cl_TR.png", width = 2000, height = 1200, res = 100) 
print(viz_cl_wl_plot)  # For ggplot objects, use print()
dev.off() 

```


##Pairs Plots
```{r Pairs Plot}
colnames(so4_cl_data_L2temp)
pairs_data <- so4_cl_data_L2temp %>% 
  select(SO4_mM, GW_salinity, Soil_EC, Soil_VWC, Water_level) %>% 
  rename("Groundwater Salinity" = GW_salinity, 
         "Soil Electrical Conductivity" = Soil_EC, 
         "Soil Volumetric Water Content" = Soil_VWC, 
         "Groundwater Water Level" = Water_level)

# , Soil_VWC, Soil_Salinity, Water_level

pairs(pairs_data, pch=19, col='orange', upper.panel=panel.smooth)

ggplot(pairs_data, aes(x = GW_salinity, y = SO4_mM)) +
  geom_point(na.rm = TRUE)+ 
  ylim(0, 45)+
  geom_smooth(method = "lm")+
  theme_classic()

# install.packages("GGally")
# install.packages("ggplot2")
library(GGally)
library(ggplot2)

ggpairs(pairs_data,
        lower = list(continuous = wrap("smooth", method = "lm")), # Add linear regression line to lower panels
        upper = list(continuous = wrap("cor", method = "pearson")), # Add correlation coefficients to upper panels
        diag = list(continuous = wrap("densityDiag")) # Add density plots to the diagonal
)


pairs_long <- pairs_data %>%
  pivot_longer(cols = -SO4_mM, names_to = "variable", values_to = "value")

so4_env <- ggplot(pairs_long, aes(x = value, y = SO4_mM)) +
  geom_point() +
  geom_smooth(method = "lm") + 
  # Adds the equation (y = mx + b)
  stat_regline_equation(label.y = 43, aes(label = ..eq.label..)) +
  # Adds the p-value and R^2
  stat_cor(label.y = 40, aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~"))) +
  facet_wrap(~ variable, scales = "free", nrow = 1) +
  labs(title = "Sulfate versus environmental variables",
       y = "SO4 (mM)",
       x = "Value") +
  theme_classic() + 
  ylim(0, 45)

so4_env

# Save as a PNG
png("Sulfate versus environmental variables.png", width = 3000, height = 1200, res = 300) 
print(so4_env)  # For ggplot objects, use print()
dev.off() 

```


## Barplots of SO4/Cl by Year, Site, and Zone
```{r Barplots of SO4/Cl, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}

data_plotting_so4_cl <- so4_cl_data 
# %>%
#   filter(!str_detect(Zone, "SW")) 

head(data_plotting_so4_cl)

#Order by Zone from Upland to Surface Water
data_plotting_so4_cl$Zone = factor(data_plotting_so4_cl$Zone, levels = c("UP", "SWAMP", "TR", "WC"))
data_plotting_so4_cl <- data_plotting_so4_cl[order(data_plotting_so4_cl$Zone), ]

#Order by Site from SWH, GCW, MSM, GWI order, which is oligohaline to polyhaline
data_plotting_so4_cl$Site = factor(data_plotting_so4_cl$Site, levels = c("SWH", "GCW", "MSM", "GWI"))
data_plotting_so4_cl <- data_plotting_so4_cl[order(data_plotting_so4_cl$Site), ]

#Plot samples to get a first look at concentrations (sanity check)
#group the data for plotting
data_plotting_so4_cl <- data_plotting_so4_cl %>%
  group_by(Site, Zone) %>%
  mutate(row_num = factor(row_number())) %>%  # create row_num as factor per group
  ungroup()




#Plot data and change colors based on Zone:
so4_cl_TR <- data_plotting_so4_cl %>%
  filter(str_detect(Zone, "TR")) %>%
  filter(str_detect(Depth_cm, "10"))

viz_cl_plot <- ggplot(so4_cl_TR, aes(x = as.character(Date), y = Cl_mM, fill = Zone)) +
  geom_boxplot() +
  facet_wrap( ~ Site, scales="free") +
  scale_fill_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "SWAMP" = "darkgrey", 
    "WC" = "#419973"
  )) +
  theme_classic() +
  labs(x = " ", y = "Chloride (mM)", title = "Chloride in Transitions Zones at 20 cm") +
  # theme(
  #   panel.spacing = unit(.05, "lines"), 
  #   panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) + 
  theme(text = element_text(size = 20))  + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))


viz_cl_plot

# Save as a PNG
png("Chloride in Transitions Zones at 20 cm.png", width = 2000, height = 1200, res = 100) 
print(viz_cl_plot)  # For ggplot objects, use print()
dev.off() 

viz_so4_plot <-  ggplot(so4_cl_TR, aes(x = as.character(Date), y = SO4_mM, fill = Zone)) +
  geom_boxplot() +
  facet_wrap( ~ Site, scales="free") +
  scale_fill_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "SWAMP" = "darkgrey", 
    "WC" = "#419973"
  )) +
  theme_classic() +
  labs(x = " ", y = "Sulfate (mM)", title = "Sulfate in Transitions Zones at 20 cm") +
  # theme(
  #   panel.spacing = unit(.05, "lines"), 
  #   panel.border = element_rect(color = "black", fill = NA, linewidth = 1))+ 
  theme(text = element_text(size = 20)) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))


viz_so4_plot

# Save as a PNG
png("Sulfate in Transitions Zones at 20 cm.png", width = 2000, height = 1200, res = 100) 
print(viz_so4_plot)  # For ggplot objects, use print()
dev.off() 

# viz_cl_so4_plot <- ggarrange(viz_cl_plot, viz_so4_plot, nrow=2, ncol=1)
# 
# # Save as a PNG
# png("Cl_SO4_boxplot_2022-2025.png", width = 2000, height = 1200, res = 100) 
# print(viz_cl_so4_plot)  # For ggplot objects, use print()
# dev.off() 


```

##Calculate Sulfate Depletion
```{r Calculate sulfate depletion}

##RSW is the constant molar ratio of Cl- to SO42- in surface seawater (RSW = 19.33; Bianchi 2006)
#Positive values of SulfateDep represent a net depletion of SO42-compared to the value predicted by the Cl:SO42- ratio in surface seawater, presumably due to sulfate reduction (Weston et al. 2006a). In a few cases, negative values of SulfateDep indicate an increase in SO42- compared to the theoretical amount
so4_cl_data <- so4_cl_data %>%
  mutate(so4_dep = (Cl_mM/19.33)-SO4_mM)

```


## Barplots of SO4 Depletion by Year, Site, and Zone
```{r Barplots of SO4/Cl, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}

data_plotting_so4_cl <- so4_cl_data %>%
  filter(Zone != "SW")

head(data_plotting_so4_cl)

#Order by Zone from Upland to Surface Water
data_plotting_so4_cl$Zone = factor(data_plotting_so4_cl$Zone, levels = c("UP", "SWAMP", "TR", "WC"))
data_plotting_so4_cl <- data_plotting_so4_cl[order(data_plotting_so4_cl$Zone), ]

#Order by Site from SWH, GCW, MSM, GWI order, which is oligohaline to polyhaline
data_plotting_so4_cl$Site = factor(data_plotting_so4_cl$Site, levels = c("SWH", "GCW", "MSM", "GWI"))
data_plotting_so4_cl <- data_plotting_so4_cl[order(data_plotting_so4_cl$Site), ]

#Plot samples to get a first look at concentrations (sanity check)
#group the data for plotting
data_plotting_so4_cl <- data_plotting_so4_cl %>%
  group_by(Site, Zone) %>%
  mutate(row_num = factor(row_number())) %>%  # create row_num as factor per group
  ungroup()


data_plotting_so4_cl$Date <- ifelse(data_plotting_so4_cl$Month < 10, paste0(data_plotting_so4_cl$Year, "0", data_plotting_so4_cl$Month), paste0(data_plotting_so4_cl$Year, data_plotting_so4_cl$Month))

#Plot data and change colors based on Depth:
so4_cl_TR <- data_plotting_so4_cl %>%
  filter(str_detect(Zone, "TR")) %>%
  filter(str_detect(Depth_cm, "10"))

viz_dep_plot_tr <- ggplot(so4_cl_TR, aes(x = as.character(Date), y = so4_dep, fill = Zone)) +
  geom_boxplot() +
  facet_wrap( ~ Site, scales="free") +
  scale_fill_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "SWAMP" = "darkgrey", 
    "WC" = "#419973"
  )) +
  theme_classic() +
  labs(x = " ", y = "Sulfate Depletion", title = "Sulfate Depletion in Transition Zones at 10 cm") +
  # theme(
  #   panel.spacing = unit(.05, "lines"), 
  #   panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) + 
  theme(text = element_text(size = 20))  + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

viz_dep_plot_tr

# Save as a PNG
png("SO4_dep_boxplot_2022-2025_10cm.png", width = 2000, height = 1200, res = 100)
print(viz_dep_plot_tr)  # For ggplot objects, use print()
dev.off()


#Plot data and change colors based on Zone:

#Compute ANOVA/Pairwise Stats per Facet
##Significance asterisks

stats <- data_plotting_so4_cl %>%
  filter(str_detect(Depth_cm, "10")) %>%
  group_by(Site) %>%
  wilcox_test(so4_dep ~ Zone) %>% # or anova_test()
  add_significance() %>%
  add_xy_position(x = "Zone") # Automate p-value positions

viz_dep_plot <- ggplot(data_plotting_so4_cl, aes(x = Zone, y = so4_dep, fill = Zone)) +
  geom_boxplot() +
  facet_wrap( ~ Site, scales="free") +
  scale_fill_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "SWAMP" = "darkgrey", 
    "WC" = "#419973"
  )) +
  theme_classic() +
  labs(x = " ", y = "Sulfate Depletion at 10 cm", title = "") +
  # theme(
  #   panel.spacing = unit(.05, "lines"), 
  #   panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) + 
  theme(text = element_text(size = 20)) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) 
  # + stat_pvalue_manual(stats, label = "p.adj.signif", tip.length = 0.01, inherit.aes = FALSE)

viz_dep_plot

#Significance letters on averaged data
# df_summ <- data_plotting_so4_cl %>%
#   filter(str_detect(Depth_cm, "10")) %>%
#   group_by(Site, Zone) %>%
#   summarise(
#     Mean = mean(so4_dep),
#     SD = sd(so4_dep),
#     .groups = 'drop'
#   ) %>%
#   # Merge the letters back into the summarized data
#   left_join(stats, by = c("Site", "Zone"))

# 
# viz_dep_plot_avg <- ggplot(df_summ, aes(x = Zone, y = Mean, , color = Zone, group = Zone)) +
#   geom_point(size = 3) +
#   geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD), width = 0.2) +
#   facet_wrap( ~ Site, scales="free") +
#   scale_color_manual(values = c(
#     "UP" = "#20063B",
#     "TR" = "#FFBC42",
#     "SWAMP" = "darkgrey",
#     "WC" = "#419973"
#   )) +
#   theme_classic() +
#   labs(x = " ", y = "Sulfate Depletion", title = "Sulfate Depletion in Transitions Zones at 20 cm") +
#   theme(text = element_text(size = 20)) +
#   theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
#   geom_text(aes(label = Letters, y = Mean + SD + 0.5),
#             position = position_dodge(0.9), vjust = 0)
# 
# viz_dep_plot_avg



##Sulfate depletion averages through time
df_summ_date <- data_plotting_so4_cl %>%
  group_by(Month, Year, Site, Zone) %>%
   dplyr::summarize(mean_SO4_dep = mean(so4_dep, na.rm=TRUE), 
            sd_SO4_dep = sd(so4_dep, na.rm=TRUE), 
            Date = first(Date))

viz_dep_plot_dates <- ggplot(df_summ_date, aes(x = Date, y = mean_SO4_dep, color = Zone, group = Zone)) +
  geom_point(size = 3) + 
  geom_errorbar(aes(ymin = mean_SO4_dep - sd_SO4_dep, ymax = mean_SO4_dep + sd_SO4_dep), width = 0.2) + 
  facet_wrap( ~ Site, scales="free") + 
  scale_color_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "SWAMP" = "darkgrey", 
    "WC" = "#419973"
  )) +
  theme_classic() +
  labs(x = " ", y = "Sulfate Depletion", title = "Sulfate Depletion in Transitions Zones at 20 cm") +
  theme(text = element_text(size = 20)) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) 

viz_dep_plot_dates


```


## Scatterplots of SO4/Cl through Time
```{r Scatterplots of SO4/Cl, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}

##Make a Full Date
data_plotting_so4_cl$Date_full <- ifelse(data_plotting_so4_cl$Month < 10, paste0(data_plotting_so4_cl$Year, "-", "0", data_plotting_so4_cl$Month), paste0(data_plotting_so4_cl$Year, "-", data_plotting_so4_cl$Month))

data_plotting_so4_cl$Date_full <- ifelse(data_plotting_so4_cl$Day < 10, paste0(data_plotting_so4_cl$Date_full, "-", "0", data_plotting_so4_cl$Day), paste0(data_plotting_so4_cl$Date_full, "-", data_plotting_so4_cl$Day))

data_plotting_so4_cl$Date_full <- as.Date(data_plotting_so4_cl$Date_full)

so4_cl_TR <- data_plotting_so4_cl %>% 
  filter(Zone == "TR")

#Plot Cl through time, colored by Zone
viz_cl_plot <- ggplot(data_plotting_so4_cl, aes(x = Date_full, y = Cl_mM, color = Zone)) +
  geom_point() +
  # geom_smooth(se = FALSE) + 
  facet_wrap( ~ Site, scales="free") +
  scale_color_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "SWAMP" = "darkgrey", 
    "WC" = "#419973"
  )) +
  theme_classic() +
  labs(x = " ", y = "Chloride (mM)", title = "Chloride throughout Time") +
  # theme(
  #   panel.spacing = unit(.05, "lines"), 
  #   panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) + 
  theme(text = element_text(size = 20))  + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))


viz_cl_plot

# Save as a PNG
png("Chloride through Time.png", width = 2000, height = 1200, res = 150)
print(viz_cl_plot)  # For ggplot objects, use print()
dev.off()



#Plot Cl through time, colored by Zone
viz_so4_plot <- ggplot(so4_cl_TR, aes(x = Date_full, y = SO4_mM, color = Zone)) +
  geom_point() +
  # geom_smooth(se = FALSE) + 
  facet_wrap( ~ Site, scales="free") +
  scale_color_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "SWAMP" = "darkgrey", 
    "WC" = "#419973"
  )) +
  theme_classic() +
  labs(x = " ", y = "Sulfate (mM)", title = "Sulfate throughout Time") +
  # theme(
  #   panel.spacing = unit(.05, "lines"), 
  #   panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) + 
  theme(text = element_text(size = 20))  + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

viz_so4_plot

# Save as a PNG
png("Sulfate through Time TR.png", width = 2000, height = 1200, res = 150)
print(viz_so4_plot)  # For ggplot objects, use print()
dev.off()


```


## Barplots of DOC by Year, Site, and Zone
```{r Barplots of DOC, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}

data_plotting_doc <- doc_data %>%
  filter(!str_detect(Zone, "SW")) 

head(data_plotting_doc)

#Order by Zone from Upland to Surface Water
data_plotting_doc$Zone = factor(data_plotting_doc$Zone, levels = c("UP", "SWAMP", "TR", "WC"))
data_plotting_doc <- data_plotting_doc[order(data_plotting_doc$Zone), ]

#Order by Site from SWH, GCW, MSM, GWI order, which is oligohaline to polyhaline
data_plotting_doc$Site = factor(data_plotting_doc$Site, levels = c("SWH", "GCW", "MSM", "GWI"))
data_plotting_doc <- data_plotting_doc[order(data_plotting_doc$Site), ]

#Plot samples to get a first look at concentrations (sanity check)
#group the data for plotting
data_plotting_doc <- data_plotting_doc %>%
  group_by(Site, Zone) %>%
  mutate(row_num = factor(row_number())) %>%  # create row_num as factor per group
  ungroup()


data_plotting_doc$Date <- ifelse(data_plotting_doc$Month < 10, paste0(data_plotting_doc$Year, "0", data_plotting_doc$Month), paste0(data_plotting_doc$Year, data_plotting_doc$Month))


#Plot data and change colors based on Zone:
doc_TR <- data_plotting_doc %>%
  filter(str_detect(Zone, "TR")) %>%
  filter(str_detect(Depth_cm, "45"))

viz_tdn_plot <- ggplot(doc_TR, aes(x = as.character(Date), y = tdn_mgL, fill = Zone)) +
  geom_boxplot() +
  facet_wrap( ~ Site, scales="free") +
  scale_fill_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "SWAMP" = "darkgrey", 
    "WC" = "#419973"
  )) +
  theme_classic() +
  labs(x = " ", y = "Chloride (mM)", title = "Chloride in Transitions Zones at 20 cm") +
  # theme(
  #   panel.spacing = unit(.05, "lines"), 
  #   panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) + 
  theme(text = element_text(size = 20))  + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))


viz_tdn_plot

# Save as a PNG
png("TDN in Transitions Zones at 20 cm.png", width = 2000, height = 1200, res = 100) 
print(viz_cl_plot)  # For ggplot objects, use print()
dev.off() 

# viz_so4_plot <-  ggplot(doc_TR, aes(x = as.character(Date), y = SO4_mM, fill = Zone)) +
#   geom_boxplot() +
#   facet_wrap( ~ Site, scales="free") +
#   scale_fill_manual(values = c(
#     "UP" = "#20063B",
#     "TR" = "#FFBC42",
#     "SWAMP" = "darkgrey", 
#     "WC" = "#419973"
#   )) +
#   theme_classic() +
#   labs(x = " ", y = "Sulfate (mM)", title = "Chloride in Transitions Zones at 20 cm") +
#   # theme(
#   #   panel.spacing = unit(.05, "lines"), 
#   #   panel.border = element_rect(color = "black", fill = NA, linewidth = 1))+ 
#   theme(text = element_text(size = 20)) + 
#   theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
# 
# 
# viz_so4_plot

# Save as a PNG
# png("Sulfate in Transitions Zones at 20 cm.png", width = 2000, height = 1200, res = 100) 
# print(viz_so4_plot)  # For ggplot objects, use print()
# dev.off() 

# viz_cl_so4_plot <- ggarrange(viz_cl_plot, viz_so4_plot, nrow=2, ncol=1)
# 
# # Save as a PNG
# png("Cl_SO4_boxplot_2022-2025.png", width = 2000, height = 1200, res = 100) 
# print(viz_cl_so4_plot)  # For ggplot objects, use print()
# dev.off() 


```

## Summarized data for Site and Zone 
```{r Summarize and Visualize Data, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}

##Boxplot Summarize SO4 and Cl by Date, Site, and Zone
viz_cl_boxplot <- ggplot(data_plotting_so4_cl, aes(x = Date,  y = Cl_ppm, fill = Zone)) +
  geom_boxplot() + 
  facet_wrap( ~ Site, scales="free_y") +
  # scale_x_discrete(drop = TRUE) +
   scale_color_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "SWAMP" = "darkgrey",
    "WC" = "#419973",
    "SW" = "#25ABE6"
  )) +
  theme_classic() +
  labs(x = " ", y = "Chloride (ppm)", title = "Porewater: Chloride") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ 
  theme(text = element_text(size = 30))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))+ 
  theme(legend.position = "none")

viz_cl_boxplot

##Summarize SO4 and Cl by Date, Site, and Zone
Dionex_avg <- all_dat_dates %>%
  group_by(Month, Year, Site, Zone) %>%
   dplyr::summarize(mean_SO4 = mean(SO4_mM, na.rm=TRUE), 
            sd_SO4 = sd(SO4_mM, na.rm=TRUE), 
            mean_Cl = mean(Cl_mM, na.rm=TRUE), 
            sd_Cl = sd(Cl_mM, na.rm=TRUE), 
            YearMonth = first(YearMonth))

# Concatenate with a day (e.g., "01" for the first day)
Dionex_avg$Date <- paste0(Dionex_avg$YearMonth, "-01")

# Convert to Date object
Dionex_avg$Date <- as.Date(Dionex_avg$Date)

# Define the specific date breaks you want to display
specific_breaks <- as.Date(c("2022-01-01", "2023-01-01", "2024-01-01", "2025-01-01"))

viz_cl_plot_avgs <- ggplot(Dionex_avg, aes(x = Date, y = mean_Cl, color = Zone, group = Zone)) +
  geom_point(size = 3) + 
  geom_errorbar(aes(ymin = mean_Cl - sd_Cl, ymax = mean_Cl + sd_Cl), width = 0.2) + 
  geom_line() + 
  facet_grid( ~ Site, scales="free_y") +
  scale_color_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "SWAMP" = "darkgrey",
    "WC" = "#419973",
    "SW" = "#25ABE6"
  )) +
  theme_classic() +
  labs(x = "Date", y = "Chloride (mM)", title = "Samples: Chloride") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  # theme(axis.text.x = element_text(angle = 90, hjust = 0.5))+ 
  theme(text = element_text(size = 30))+
  scale_x_date(breaks = specific_breaks, date_labels = "%Y", expand = expansion(mult = c(0.18, 0.05)))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))

# viz_cl_plot_avgs


viz_so4_plot_avgs <- ggplot(Dionex_avg, aes(x = Date, y = mean_SO4, color = Zone, group = Zone)) +
  geom_point(size = 3) + 
  geom_errorbar(aes(ymin = mean_SO4 - sd_SO4, ymax = mean_SO4 + sd_SO4), width = 0.2) + 
  geom_line() + 
  facet_grid( ~ Site, scales="free_y") +
  scale_color_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "SWAMP" = "darkgrey",
    "WC" = "#419973",
    "SW" = "#25ABE6"
  )) +
  theme_classic() +
  labs(x = " ", y = "Sulfate (mM)", title = "Samples: Sulfate") +
  theme(
    panel.spacing = unit(0.5, "lines"), 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) +
  # theme(axis.text.x = element_text(angle = 90, hjust = 0.5))+ 
  theme(text = element_text(size = 30))+
  scale_x_date(breaks = specific_breaks, date_labels = "%Y", expand = expansion(mult = c(0.18, 0.05)))+
  theme(axis.ticks.x = element_line(linewidth = 1))+
  theme(axis.ticks.length.x = unit(0.25, "cm"))

# viz_so4_plot_avgs

ggarrange(viz_cl_plot, viz_so4_plot, nrow=2, ncol=1)

```

##Modelling
```{r}
##ANOVA comparison of sulfate across years within each zone
data_plotting_so4_cl$Year <- as.factor(data_plotting_so4_cl$Year)

results <- data_plotting_so4_cl %>%
  group_by(Site, Zone) %>%
  do(tidy(aov(SO4_mM ~ Year, data = .)))

##Separate ANOVA by Site and Zone
so4_cl_GCW_TR <- so4_cl_data %>%
  filter(str_detect(Site, "GCW")) %>%
  filter(str_detect(Zone, "TR")) 


library(lme4)
library(lmerTest)

so4_cl_data$SampleUnit <- interaction(
                             so4_cl_data$Site,
                             so4_cl_data$Zone,
                             so4_cl_data$Depth_cm,
                             so4_cl_data$Replicate)

model <- lmer(SO4_mM ~ Year * Month + (1 | SampleUnit), data = so4_cl_data)

anova(model)

# Type III Analysis of Variance Table with Satterthwaite's method
#            Sum Sq Mean Sq NumDF  DenDF F value  Pr(>F)  
# Year         0.49    0.49     1 2146.4  0.0132 0.90853  
# Month      184.12  184.12     1 2157.5  4.9645 0.02598 *
# Year:Month 181.44  181.44     1 2157.5  4.8924 0.02708 *


# The ANOVA table will show:
# Year → overall annual differences
# Month → seasonal differences
# Year:Month → does seasonality differ among years?
# That interaction is often ecologically very important.


plot(model)
qqnorm(residuals(model))
qqline(residuals(model))


so4_cl_data$log_SO4 <- log(so4_cl_data$SO4_mM + 0.01)

model <- lmer(log_SO4 ~ Year * Month + (1 | SampleUnit), data = so4_cl_data)

anova(model)

plot(model)
qqnorm(residuals(model))
qqline(residuals(model))

sample_sizes <- so4_cl_data %>%
  group_by(Site, Zone, Depth_cm, Replicate) %>%
  summarise(n = n(), .groups = "drop")

sample_sizes

table(sample_sizes$n)
```

## Barplots of H2S by Year, Site, and Zone
```{r Barplots of H2S, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}

data_plotting_h2s <- h2s_data %>%
  filter(!str_detect(Zone, "SW"))

data_plotting_h2s <- data_plotting_h2s %>% filter(!is.na(Year))

#Order by Zone from Upland to Surface Water
data_plotting_h2s$Zone = factor(data_plotting_h2s$Zone, levels = c("UP", "SWAMP", "TR", "WC"))
data_plotting_h2s <- data_plotting_h2s[order(data_plotting_h2s$Zone), ]

#Order by Site from SWH, GCW, MSM, GWI order, which is oligohaline to polyhaline
data_plotting_h2s$Site = factor(data_plotting_h2s$Site, levels = c("SWH", "GCW", "MSM", "GWI"))
data_plotting_h2s <- data_plotting_h2s[order(data_plotting_h2s$Site), ]

# #Order by Year from 2022-2024
# data_plotting$Site = factor(data_plotting$Site, levels = c("SWH", "GCW", "MSM", "GWI"))
# data_plotting <- data_plotting[order(data_plotting$Site), ]

#Plot samples to get a first look at concentrations (sanity check)
#group the data for plotting
data_plotting_h2s <- data_plotting_h2s %>%
  group_by(Site, Zone) %>%
  mutate(row_num = factor(row_number())) %>%  # create row_num as factor per group
  ungroup()

data_plotting_h2s$Date <- as.Date(paste(data_plotting_h2s$Year, data_plotting_h2s$Month, data_plotting_h2s$Day, sep = "-"))


#Plot data and change colors based on Zone:
viz_H2S_plot <- ggplot(data_plotting_h2s, aes(x = as.character(Year), y = H2S_mean_uM, fill = Zone)) +
  geom_boxplot() +
  facet_wrap( ~ Site, scales="free", nrow = 1) +
  scale_fill_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "SWAMP" = "darkgrey",
    "WC" = "#419973"
  )) +
  theme_classic() +
  labs(x = " ", y = "Sulfide (uM)", title = "Samples: Sulfide") +
  # theme(
  #   panel.spacing = unit(.05, "lines"), 
  #   panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) + 
  theme(text = element_text(size = 20))

viz_H2S_plot

# Save as a PNG
png("H2S_boxplot_2022-2025.png", width = 2000, height = 600, res = 100) 
print(viz_H2S_plot)  # For ggplot objects, use print()
dev.off() 

```

## Barplots of DOC by Year, Site, and Zone
```{r Barplots of DOC, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}

data_plotting_doc <- doc_data %>%
  filter(!str_detect(Zone, "SW"))

#Order by Zone from Upland to Surface Water
data_plotting_doc$Zone = factor(data_plotting_doc$Zone, levels = c("UP", "SWAMP", "TR", "WC", "SW"))
data_plotting_doc <- data_plotting_doc[order(data_plotting_doc$Zone), ]

#Order by Site from SWH, GCW, MSM, GWI order, which is oligohaline to polyhaline
data_plotting_doc$Site = factor(data_plotting_doc$Site, levels = c("SWH", "GCW", "MSM", "GWI"))
data_plotting_doc <- data_plotting_doc[order(data_plotting_doc$Site), ]

#Plot samples to get a first look at concentrations (sanity check)
#group the data for plotting
data_plotting_doc <- data_plotting_doc %>%
  group_by(Site, Zone) %>%
  mutate(row_num = factor(row_number())) %>%  # create row_num as factor per group
  ungroup()

data_plotting_doc$Date <- as.Date(paste(data_plotting_doc$Year, data_plotting_doc$Month, data_plotting_doc$Day, sep = "-"))


#Plot data and change colors based on Zone:
viz_NPOC_plot <- ggplot(data_plotting_doc, aes(x = as.character(Year), y = npoc_mgL, fill = Zone)) +
  geom_boxplot() +
  facet_wrap( ~ Site, scales="free", nrow = 1) +
  scale_fill_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "SWAMP" = "darkgrey",
    "WC" = "#419973",
    "SW" = "#25ABE6"
  )) +
  theme_classic() +
  labs(x = " ", y = "NPOC (mg/L)", title = "Barplot of NPOC by Year, Site and Zone") +
  # theme(
  #   panel.spacing = unit(.05, "lines"), 
  #   panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) + 
  theme(text = element_text(size = 30))

 viz_NPOC_plot
 
 viz_TDN_plot <- ggplot(data_plotting_doc, aes(x = as.character(Year), y = tdn_mgL, fill = Zone)) +
  geom_boxplot() +
  facet_wrap( ~ Site, scales="free", nrow = 1) +
  scale_fill_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "SWAMP" = "darkgrey",
    "WC" = "#419973",
    "SW" = "#25ABE6"
  )) +
  theme_classic() +
  labs(x = " ", y = "TDN (mg/L)", title = "Barplot of TDN by Year, Site and Zone") +
  # theme(
  #   panel.spacing = unit(.05, "lines"), 
  #   panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) + 
  theme(text = element_text(size = 30))

 viz_TDN_plot

 viz_doc_plot <- ggarrange(viz_NPOC_plot, viz_TDN_plot, nrow=2, ncol=1)

 
# Save as a PNG
png("DOC_boxplot_2022-2025.png", width = 2000, height = 1200, res = 100) 
print(viz_doc_plot)  # For ggplot objects, use print()
dev.off() 

```

##Combine SO4/Cl and DOC data
```{r}

doc_so4_data <- merge(doc_data, so4_cl_data, by = c("Year", "Month", "Site", "Zone", "Replicate", "Depth_cm"))

```

##Plot DOC vs SO4/Cl
```{r}

doc_so4_data_TR <- doc_so4_data %>%
  filter(str_detect(Zone, "TR"))

##Different lines for each zone (UP, TR, WC)
viz_doc_so4_plot <- ggplot(doc_so4_data_TR, aes(x = SO4_ppm, y = npoc_mgL)) +
  geom_point(aes(color = Zone), size = 3) +
  stat_poly_eq(use_label(c("R2", "P")), size = 6) + 
  stat_poly_line(formula = y ~ x, color = "black")+   facet_wrap( ~ Site, scales="free", nrow = 1) +
  scale_color_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "SWAMP" = "darkgrey",
    "WC" = "#419973",
    "SW" = "#25ABE6"
  )) +
  theme_classic() +
  labs(x = "SO4 (ppm)", y = "NPOC (mg/L)", title = "Plot of NPOC vs SO4 for TR Zone") +
  # theme(
  #   panel.spacing = unit(.05, "lines"), 
  #   panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) + 
  theme(text = element_text(size = 30)) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

 viz_doc_so4_plot
 

##Different lines for each zone (UP, TR, WC)
viz_doc_cl_plot <- ggplot(doc_so4_data_TR, aes(x = Cl_ppm, y = npoc_mgL)) +
  geom_point(aes(color = Zone), size = 3) +
  stat_poly_eq(use_label(c("R2", "P")), size = 6) + 
  stat_poly_line(formula = y ~ x, color = "black")+ 
  facet_wrap( ~ Site, scales="free", nrow = 1) +
  scale_color_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "SWAMP" = "darkgrey",
    "WC" = "#419973",
    "SW" = "#25ABE6"
  )) +
  theme_classic() +
  labs(x = "Cl (ppm)", y = "NPOC (mg/L)", title = "Plot of NPOC vs Cl for TR Zone") +
  # theme(
  #   panel.spacing = unit(.05, "lines"), 
  #   panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) + 
  theme(text = element_text(size = 30)) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

viz_doc_cl_plot


# Save as a PNG
png("DOC_versus_SO4_TR.png", width = 2000, height = 1200, res = 100) 
print(viz_doc_so4_plot)  # For ggplot objects, use print()
dev.off() 

# Save as a PNG
png("DOC_versus_Cl_TR.png", width = 2000, height = 1200, res = 100) 
print(viz_doc_cl_plot)  # For ggplot objects, use print()
dev.off() 

```


##Combine SO4/Cl and H2S data
```{r}

h2s_so4_data <- merge(h2s_data, so4_cl_data, by = c("Year", "Month", "Site", "Zone", "Replicate", "Depth_cm"))

```

##Plot SO4/Cl vs H2S
```{r}

h2s_so4_data_TR <- h2s_so4_data %>%
  filter(str_detect(Zone, "TR"))

##Different lines for each zone (UP, TR, WC)
viz_h2s_so4_plot <- ggplot(h2s_so4_data_TR, aes(x = H2S_mean_uM, y = SO4_ppm)) +
  geom_point(aes(color = Zone), size = 3) +
  stat_poly_eq(use_label(c("R2", "P")), size = 6) + 
  stat_poly_line(formula = y ~ x, color = "black")+   facet_wrap( ~ Site, scales="free", nrow = 1) +
  scale_color_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "SWAMP" = "darkgrey",
    "WC" = "#419973",
    "SW" = "#25ABE6"
  )) +
  theme_classic() +
  labs(x = "H2S (uM)", y = "SO4 (mM)", title = "Plot of H2S vs SO4 for TR Zone") +
  # theme(
  #   panel.spacing = unit(.05, "lines"), 
  #   panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) + 
  theme(text = element_text(size = 30)) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

 viz_h2s_so4_plot


# Save as a PNG
png("H2S vs SO4 for TR Zone.png", width = 2000, height = 1200, res = 100) 
print(viz_h2s_so4_plot)  # For ggplot objects, use print()
dev.off() 

```

## Scatterplot of Sulfate Depletion by Depth
```{r Scatterplot of SO4 Depletion by Depth, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}

##Sulfate depletion averages through time
df_summ_date <- data_plotting_so4_cl %>%
  filter(Zone == "TR") %>% 
  group_by(Depth_cm, Site, Zone) %>%
   dplyr::summarize(mean_SO4_dep = mean(so4_dep, na.rm=TRUE), 
            sd_SO4_dep = sd(so4_dep, na.rm=TRUE), 
            Date = first(Date))

viz_dep_plot_depths <- ggplot(df_summ_date, aes(x = Depth_cm, y = mean_SO4_dep, color = Zone, group = Zone)) +
  geom_point(size = 3) + 
  geom_errorbar(aes(ymin = mean_SO4_dep - sd_SO4_dep, ymax = mean_SO4_dep + sd_SO4_dep), width = 0.2) + 
  facet_wrap( ~ Site, scales="free") + 
  scale_color_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "SWAMP" = "darkgrey", 
    "WC" = "#419973"
  )) +
  theme_classic() +
  labs(x = " ", y = "Sulfate Depletion", title = "Sulfate Depletion in Transitions Zones at 20 cm") +
  theme(text = element_text(size = 20)) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) 

viz_dep_plot_depths

#Plot data and change colors based on Zone:
viz_so4_dep_depth_scatter <-  ggplot(so4_data_TR, aes(x = Depth_cm, y = so4_dep)) +
  geom_point(aes(color = Zone), size = 3) +
  stat_poly_eq(use_label(c("R2", "P")), size = 6) + 
  stat_poly_line(formula = y ~ x, color = "black")+   facet_wrap( ~ Site, scales="free", nrow = 1) +
  scale_color_manual(values = c(
    "UP" = "#20063B",
    "TR" = "#FFBC42",
    "SWAMP" = "darkgrey",
    "WC" = "#419973",
    "SW" = "#25ABE6"
  )) +
  theme_classic() +
  labs(x = "Sampling Depth (cm)", y = "Sulfate Depletion", title = "") +
  # theme(
  #   panel.spacing = unit(.05, "lines"), 
  #   panel.border = element_rect(color = "black", fill = NA, linewidth = 1))+ 
  theme(text = element_text(size = 30))+ 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

viz_so4_dep_depth_scatter

# Save as a PNG
png("SO4_dep_by_depth.png", width = 2000, height = 1200, res = 100) 
print(viz_so4_dep_plot_depth)  # For ggplot objects, use print()
dev.off() 

```

## Barplot of Sulfate Depletion by Depth
```{r Barplots of SO4 Depletion by Depth, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}

#Plot data and change colors based on Zone:
viz_so4_dep_plot_depth <-  ggplot(data_plotting_so4_cl, aes(x = Zone, y = so4_dep, fill = as.character(Depth_cm))) +
  geom_boxplot() +
  facet_wrap( ~ Site, scales="free", nrow = 1) +
  scale_fill_manual(values = c(
    "10" = "lightblue",
    "20" = "blue",
    "45" = "darkblue"
  )) +
  theme_classic() +
  labs(x = "Sampling Depth (cm)", y = "Sulfate Depletion", title = "") +
  # theme(
  #   panel.spacing = unit(.05, "lines"), 
  #   panel.border = element_rect(color = "black", fill = NA, linewidth = 1))+ 
  theme(text = element_text(size = 30))+ 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

viz_so4_dep_plot_depth

# Save as a PNG
png("SO4_dep_by_depth.png", width = 2000, height = 1200, res = 100) 
print(viz_so4_dep_plot_depth)  # For ggplot objects, use print()
dev.off() 

```

## Barplots of SO4/Cl by Depth
```{r Barplots of SO4/Cl by Depth, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}

data_plotting_so4_cl <- data_plotting_so4_cl %>% 
  filter(Zone != "SW")

#Plot data and change colors based on Zone:
# viz_cl_plot_depth <- ggplot(data_plotting_so4_cl, aes(x = Zone_1, y = Cl_mM, fill = as.character(Depth_cm))) +
#   geom_boxplot() +
#   facet_wrap( ~ Site, scales="free", nrow = 1) +
#   scale_fill_manual(values = c(
#     "10" = "lightblue",
#     "20" = "blue",
#     "45" = "darkblue"
#   )) +
#   theme_classic() +
#   labs(x = "Sampling Depth (cm)", y = "Chloride (mM)", title = "Samples: Chloride") +
#   # theme(
#   #   panel.spacing = unit(.05, "lines"), 
#   #   panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) + 
#   theme(text = element_text(size = 30))
# 
# viz_cl_plot_depth


##Calculate average and sd
so4_cl_avg <- data_plotting_so4_cl %>%
  group_by(Site_1, Zone_1, Depth_cm) %>%
   dplyr::summarize(mean_SO4 = mean(SO4_mM, na.rm=TRUE), 
            sd_SO4 = sd(SO4_mM, na.rm=TRUE))


viz_so4_plot_depth <- ggplot(so4_cl_avg, aes(x = as.numeric(Depth_cm), y = mean_SO4, color = Zone_1, group = Zone_1)) +
  geom_point(size = 3) + 
  geom_errorbar(aes(ymin = mean_SO4 - sd_SO4, ymax = mean_SO4 + sd_SO4), width = 0.2) + 
  geom_line() + 
  facet_wrap( ~ Site_1, scales="free") + 
  scale_fill_manual(values = c(
    "Upland" = "#20063B",
    "Transition" = "#FFBC42",
    "Swamp" = "darkgrey",
    "Wetland" = "#419973",
    "Surface Water" = "#25ABE6"
  )) + 
  theme_classic() +
  labs(x = "Sampling Depth (cm)", y = "Sulfate (mM)", title = "Sulfate by Depth") +
  theme(text = element_text(size = 20)) + 
  coord_flip() + 
  scale_x_continuous(trans = "reverse")

viz_so4_plot_depth


##Anova
# Model: Sulfate explained by Depth, Site, and Zone, including interactions
# * indicates all main effects and interaction terms
anova_model <- aov(SO4_mM ~ Depth_cm * Site * Zone, data = data_plotting_so4_cl)

# Display results
summary(anova_model)

#check normality
plot(anova_model, 2)

stats <- data_plotting_so4_cl %>%
  group_by(Site, Zone) %>%
  wilcox_test(SO4_mM ~ Depth_cm) %>% # or anova_test()
  add_significance() %>%
  add_xy_position(x = "Zone") # Automate p-value positions

# viz_so4_plot_depth <-  ggplot(data_plotting_so4_cl, aes(x = Depth_cm, y = SO4_mM, fill = Zone_1)) +
#   geom_boxplot() +
#   facet_wrap( ~ Site, scales="free", nrow = 1) +
#   scale_fill_manual(values = c(
#     "Upland" = "#20063B",
#     "Transition" = "#FFBC42",
#     "Swamp" = "darkgrey",
#     "Wetland" = "#419973",
#     "Surface Water" = "#25ABE6"
#   )) +
#   theme_classic() +
#   labs(x = "Sampling Depth (cm)", y = "Sulfate (mM)", title = "Sulfate by Depth") +
#   theme(text = element_text(size = 30)) +
#   labs(fill = "Depth") +
#   coord_flip()
# 
# viz_so4_plot_depth

# Save as a PNG
png("SO4_by_depth_avgs.png", width = 2000, height = 1000, res = 100) 
print(viz_so4_plot_depth)  # For ggplot objects, use print()
dev.off() 

# viz_cl_so4_plot_depth <- ggarrange(viz_cl_plot_depth, viz_so4_plot_depth, nrow=2, ncol=1)

# # Save as a PNG
# png("Cl_SO4_by_depth.png", width = 2000, height = 1200, res = 100) 
# print(viz_cl_so4_plot_depth)  # For ggplot objects, use print()
# dev.off() 

```
## Barplots of H2S by Depth
```{r Barplots of H2S by Depth, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}

#Plot data and change colors based on Zone:
viz_H2S_plot_depth <- ggplot(data_plotting_h2s, aes(x = Zone, y = H2S_mean_uM, fill = as.character(Depth_cm))) +
  geom_boxplot() +
  facet_wrap( ~ Site, scales="free", nrow = 1) +
  scale_fill_manual(values = c(
  "10" = "lightblue",
    "20" = "blue",
    "45" = "darkblue" )) +
  theme_classic() +
  labs(x = "Sampling Depth (cm)", y = "Sulfide (uM)", title = "Samples: Sulfide") +
  # theme(
  #   panel.spacing = unit(.05, "lines"), 
  #   panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) + 
  theme(text = element_text(size = 30))

viz_H2S_plot_depth

# Save as a PNG
png("H2S_boxplot_by_depth.png", width = 2000, height = 600, res = 100) 
print(viz_H2S_plot_depth)  # For ggplot objects, use print()
dev.off() 

```

## Barplots of DOC by Depth
```{r Barplots of DOC by Depth, echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=10, dpi = 300}

#Plot data and change colors based on Zone:
viz_NPOC_plot_depth <- ggplot(data_plotting_doc, aes(x = Zone, y = npoc_mgL, fill = as.character(Depth_cm))) +
  geom_boxplot() +
  facet_wrap( ~ Site, scales="free", nrow = 1) +
  scale_fill_manual(values = c(
    "10" = "lightblue",
    "20" = "blue",
    "45" = "darkblue"
  )) +
  theme_classic() +
  labs(x = "Sampling Depth (cm)", y = "NPOC (mg/L)", title = "Samples: NPOC") +
  # theme(
  #   panel.spacing = unit(.05, "lines"), 
  #   panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) + 
  theme(text = element_text(size = 30))

 viz_NPOC_plot_depth
 
 viz_TDN_plot_depth <- ggplot(data_plotting_doc, aes(x = Zone, y = tdn_mgL, fill = as.character(Depth_cm))) +
  geom_boxplot() +
  facet_wrap( ~ Site, scales="free", nrow = 1) +
  scale_fill_manual(values = c(
   "10" = "lightblue",
    "20" = "blue",
    "45" = "darkblue"
  )) +
  theme_classic() +
  labs(x = "Sampling Depth (cm)", y = "TDN (mg/L)", title = "Samples: TDN") +
  # theme(
  #   panel.spacing = unit(.05, "lines"), 
  #   panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) + 
  theme(text = element_text(size = 30))

 viz_TDN_plot_depth

 viz_doc_plot_depth <- ggarrange(viz_NPOC_plot_depth, viz_TDN_plot_depth, nrow=2, ncol=1)

 
# Save as a PNG
png("DOC_boxplot_by_depth.png", width = 2000, height = 1200, res = 100) 
print(viz_doc_plot_depth)  # For ggplot objects, use print()
dev.off() 

```
##Plot all analytes by depth
```{r}

viz_all_analytes_depth <- ggarrange(viz_so4_dep_plot_depth, viz_so4_plot_depth, viz_cl_plot_depth, viz_H2S_plot_depth, viz_NPOC_plot_depth, viz_TDN_plot_depth, nrow=6, ncol=1)
 
# Save as a PNG
png("Analytes_boxplot_by_depth.png", width = 2000, height = 3500, res = 100) 
print(viz_all_analytes_depth)  # For ggplot objects, use print()
dev.off() 
```

##Scatterplot of SO4 and H2S by Depth
```{r}

h2s_so4_data_TR <- h2s_so4_data %>%
  filter(str_detect(Zone, "TR"))

##Different lines for each zone (UP, TR, WC)
viz_so4_plot <- ggplot(h2s_so4_data, aes()) +
  geom_boxplot(aes(x = as.character(Depth_cm), y = SO4_ppm)) +
  # stat_poly_eq(use_label(c("R2", "P")), size = 6) + 
  # stat_poly_line(formula = y ~ x, color = "black")+   facet_wrap( ~ Site, scales="free", nrow = 1) +
  # scale_color_manual(values = c(
  #   "UP" = "#20063B",
  #   "TR" = "#FFBC42",
  #   "SWAMP" = "darkgrey",
  #   "WC" = "#419973",
  #   "SW" = "#25ABE6"
  # )) +
  theme_classic() +
  labs(x = "Depth (cm)", y = "SO4 (ppm)", title = "Plot of SO4 vs Depth for TR Zone") +
  # theme(
  #   panel.spacing = unit(.05, "lines"), 
  #   panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) + 
  theme(text = element_text(size = 30)) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

 viz_so4_plot

 viz_h2s_plot <- ggplot(h2s_so4_data, aes()) +
   geom_boxplot(aes(x = as.character(Depth_cm), y = H2S_mean_uM)) +
  # stat_poly_eq(use_label(c("R2", "P")), size = 6) + 
  # stat_poly_line(formula = y ~ x, color = "black")+   facet_wrap( ~ Site, scales="free", nrow = 1) +
  # scale_color_manual(values = c(
  #   "UP" = "#20063B",
  #   "TR" = "#FFBC42",
  #   "SWAMP" = "darkgrey",
  #   "WC" = "#419973",
  #   "SW" = "#25ABE6"
  # )) +
  theme_classic() +
  labs(x = "Depth (cm)", y = "H2S (uM)", title = "Plot of H2S vs Depth for TR Zone") +
  # theme(
  #   panel.spacing = unit(.05, "lines"), 
  #   panel.border = element_rect(color = "black", fill = NA, linewidth = 1)) + 
  theme(text = element_text(size = 30)) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

 viz_h2s_plot

# Save as a PNG
png("SO4 vs Depth.png", width = 2000, height = 1200, res = 100) 
print(viz_so4_plot)  # For ggplot objects, use print()
dev.off() 

# Save as a PNG
png("H2S vs Depth.png", width = 2000, height = 1200, res = 100) 
print(viz_h2s_plot)  # For ggplot objects, use print()
dev.off() 

```
